Evaluation of +S2 sea ice forecast and some other better title

### 2025-03-07 analysis/paper/access-ice.tex
renders

\firstpage{1}
\maketitle
\begin{abstract}
The abstract goes here. It can also be on \emph{multiple lines}.
\end{abstract}
\copyrightstatement{The author's copyright for this publication is
transferred to institution/company.}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, frame hidden, sharp corners, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi

@@

The ACCESS-S2 hindcast used in this study runs for the period
1981--2018. Each forecast consists of 9 ensemble members, 3 of which run
for 279 days and 6 for 42 days. Anomalies are be taken with respect to
the 1981--2011 climatology computed from the daily ACCESS-S2 reanalysis
and smoothed with a 11 day running mean.

The ACCESS-S1 hindcast runs for the period 1990--2012, with 7-member
forecasts integrated for 217 days. No reanalysis is available so we are
not able to compute a climatology for this forecast, therefore we don't
show metrics based on anomalies.

@@

\caption{\label{fig-hindcast-extent}Median sea ice extent for al
hindcasts initialised the first of the month for ACCESS-S2 and ACCESS-S1
in colours representing the start month. In black, the median sea ice
extent of NOAA/NSIDC CDRV4.}

@@

Figure~\ref{fig-hindcast-extent} a shows the seasonal cycle of Sea Ice
Extent for the ACCESS-S2 hindcast and NOAA/NSIDC CDRV4 ACCESS-S2 shows a
severe low extent bias, especially in the late summer-early autumn. This
is due primarily to faster melt during a longer melt season between
January and March and slower growth during March and April. This is then
balanced with faster growth between May and July
(Fig.~\ref{fig-mean-growth}). Many sea ice models exhibit this
systematic underestimation during the sea ice minimum and early freezing
season \citep{massonnet2023} and could indicate problems in the
representation of thermodynamics in the model \citep{zampieri2019}.

ACCESS-S1 has a smaller low bias than ACCESS-S2, especially at low lags
(Fig.~\ref{fig-hindcast-extent}) even though the typical growth rates
are very similar between both models (Fig.~\ref{fig-mean-growth}). At
larger lags, ACCESS-S1's bias in summer and autumn is very similar to
ACCESS-S2's. This suggests that this lower sea ice state is closer to
the models' equilibrium, indicating that it is an issue with model
formulation that was being corrected by the data assimilation system in
ACCESS-S1 and not in ACCESS-S2.

At large lags, sea ice extent loses most of the initial condition memory
and reverts to the model's preferred equilibrium state. Therefore we can
estimate the latter using the hindcasts with the largest possible lag,
which is shown in triangles in Figure~\ref{fig-hindcast-extent} for the
same dates as the initial conditions. The difference between the two is
the effect of the data assimilation.

The equilibrium of ACCESS-S1 and ACCESS-S2 is very similar (comparing
the triangles in each panel in Fig.~\ref{fig-hindcast-extent}), owing to
them having the same model formulation. From June to October, in
ACCESS-S2 circles are closer to observations than to the triangles,
indicating that the information from the ocean and atmosphere data
assimilation is getting to sea ice and improving the initial conditions.
The rest of the year, there is little if any difference between circles
and triangles in ACCESS-S2, indicating that almost no data assimilation
is taking place and the sea ice component of the model is virtually
free-running. In ACCESS-S1, on the other hand, circles are very close to
observations year-round.

@@

\caption{\label{fig-rmse}Median and 95\% coverage of sea ice
concentration anomalies RMSE as a function of forecast lag for all
forecast initialised on the first of each month compared with a
reference forecast of persistence of anomalies.}

>>>>

\firstpage{1}
\maketitle
\begin{abstract}
The abstract goes here. It can also be on \emph{multiple lines}.
\end{abstract}
\copyrightstatement{The author's copyright for this publication is
transferred to institution/company.}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, boxrule=0pt, enhanced, breakable]}{\end{tcolorbox}}\fi

@@

The ACCESS-S2 hindcast used in this study runs for the period
1981--2018. Each forecast consists of 9 ensemble members built from
three consecutive 3-member forecasts initialised at the first of every
month and the two previous days and run from 279 days. The ACCESS-S1
hindcast is built in the same manner and running for the period
1990--2012.

Anomalies for each hindcast are be taken with respect to a
forecast-dependent climatology for the period 1990--2012. This serves as
a first-order correction of model drift.

@@

\caption{\label{fig-hindcast-extent}Median sea ice extent for al
hindcasts initialised the first of the month for ACCESS-S2 and ACCESS-S1
in colours representing the start month with the median sea ice extent
of NOAA/NSIDC CDRV4 in the left column and the median difference in the
right column. The NOAA/NSIDC CDRV4 climatology is computed in the period
corresponding to each hindcast. Circles represent the initial conditions
at the first of every month and triangles represent the median value at
the first of every month forecasted with the largest possible lead time.
Colours indicate the initialisation month of the forecast.}

@@

Figure~\ref{fig-hindcast-extent} a shows the median sea ice extent (left
column) and median difference with respect of NOAA/NSIDC CDRV4 for the
ACCESS-S2 and ACCESS-S1 hindcasts. Median extent of initial conditions
at the first of every month are indicated with circles, while median
extent at the same date but forecasted with the largest lead time
possible for each model (between 274 and 277 days for ACCESS-S2 and
between 213 and 216 days for ACCESS-S1). At this large lead time, the
the information of the initial conditions is essentially lost and the
forecast reverts to each model's preferred equilibrium state.

Both models' equilibrium state (triangles) show low bias of sea ice
extent, particularly in the late-autumn and winter months. This is due
primarily to faster melt during a longer melt season between January and
March and slower growth during March and April. This is then partially
balanced with faster growth between May and July
(Fig.~\ref{fig-mean-growth}). Many sea ice models exhibit this
systematic underestimation during the sea ice minimum and early freezing
season \citep{massonnet2023} and could indicate problems in the
representation of thermodynamics in the model \citep{zampieri2019}. It
is not surprising that both forecasting systems converge to a similar
climatology because they share the same model configuration.

ACCESS-S2 initial conditions (circles) also have a low extent bias,
especially in the late summer-early autumn, while ACCESS-S1 initial
conditions are very close to observations. The faster growth between May
and July reduces the low bias in ACCESS-S2 for forecasts initialised
from April to June, but it adds a high bias in ACCESS-S1.

The difference between the initial conditions and the model preferred
state can be attributed to the effect of data assimilation, which in
ACCESS-S2 is due solely to atmospheric and oceanic data assimilation.
From June to October, in ACCESS-S2 circles are closer to observations
than to the triangles, indicating that the information from the ocean
and atmosphere data assimilation is affecting sea ice and improving the
initial conditions. The rest of the year there is little if any
difference between circles and triangles in ACCESS-S2, indicating that
almost no data assimilation is taking place and the sea ice component of
the model is virtually free-running.

@@

\caption{\label{fig-rmse}Median RMSE of sea ice concentration anomalies
as a function of forecast lag for all forecast initialised on the first
of each month compared with a reference forecast of persistence of
anomalies.}
